---
title: "Statistical_Learning_Lab_1"
author: "Gutman Abigail & Shalom Shahar"
date: "22 3 2021"
output: html_document
---

```{r setup, include=FALSE}
# tools and data upload
library(ggplot2)
library(dplyr)
library(scales)   # needed for formatting y-axis labels to non-scientific type
library(tidyr)
library(tidyverse)
library(ggthemes)
library(datasets)
library(nycflights13)
library(kableExtra) #table styling
library(e1071) # kurtosis and skeeness
library(MASS) # reg modeling
library(corrplot)

```


```{r}
options(scipen = 999)
flights <- as.data.frame(nycflights13::flights)
wheather <- nycflights13::weather
airports <- nycflights13::airports
planes <- nycflights13::planes
```

```{r, echo=FALSE, results='asis',fig.keep='all'}
summary(flights)
flights_con <- flights[,c(2:9, 11 ,15:19)]
flights_catgory <- flights[,c(10 ,12:15)]

catgory <- colnames(flights_catgory) 
carrier <- flights_catgory %>% count(carrier)# helpful
kable(carrier, row.names = F) %>% kable_styling()

tailnum <-flights_catgory %>% count(tailnum)
kable(tailnum, row.names = F) %>% kable_styling()

origin <- flights_catgory %>% count(origin) #helpful
kable(origin, row.names = F) %>% kable_styling()

dest <- flights_catgory %>% count(dest) # need to check
kable(dest, row.names = F) %>% kable_styling()

air_time <- flights_catgory %>% count(air_time)#need to check
kable(air_time, row.names = F) %>% kable_styling()

countinuous <- colnames(flights_con)
ggplot(data = flights_con) +
  geom_bar(mapping = aes(x = month))  # 2 חשוד

ggplot(data = flights_con) +
  geom_bar(mapping = aes(x = day))  # 30 לחודש חשוד

ggplot(data = flights, mapping = aes(x = dep_time/100)) + 
  geom_histogram(binwidth = 0.25) # טיסות שממריאות בלילה ?

ggplot(data = flights, mapping = aes(x = arr_time/100)) + 
  geom_histogram(binwidth = 0.25)  # טיסות שמגיעות בלילה?


ggplot(data = flights, mapping = aes(x = sched_dep_time/100)) + 
  geom_histogram(binwidth = 0.25)

ggplot(data = flights, mapping = aes(x = dep_delay)) + 
  geom_histogram(binwidth = 1) + xlim(0,300)
dep_delay_over <- flights %>% filter(dep_delay > 300)


ggplot(data = flights, mapping = aes(x = arr_delay)) + 
  geom_histogram(binwidth = 1) + xlim(0,300)
arr_delay_over <- flights %>% filter(arr_delay > 300)


ggplot(data = flights, mapping = aes(x = sched_arr_time/100)) + 
  geom_histogram(binwidth = 0.25)

ggplot(data = flights, mapping = aes(x = flight)) + 
  geom_histogram(binwidth = 1) # מזהה טיסה

ggplot(data = flights, mapping = aes(x = distance)) + 
  geom_histogram(binwidth = 1) + xlim (0,2000)
long_dist <- flights %>% filter(distance > 2000)

```


1 Graph Critique
flight figure 1: the prcantge of delayes in departures,only obove 15 minutes,flights out of DEN airport.
Trying to understande in which flights routes coming out of DEN, have more delyes.
flight figure 2:
The reltionship between number of the delyed flights to the total number of flights. 
צריך להמשיך לזיין את השכל פה, צריך לנהל דיון קוקי





2.
```{r}



```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
